---
title: "Example Public Regional Report Card: Coast-wide Bull Kelp"
# These can be turned on to have different inputs to the RMarkdown file
output:
  html_document:
    theme: yeti
    css: "style.css"
    # toc: yes
# author: Mairin Deith
# date: 28-02-2025
---

<!-- Remove the left padding -->
<style>
.main-container {
    max-width: 1080px;
    margin-left: 0;
    margin-right: auto;
}
</style>


## Coast-wide kelp at-a-glance

**Data shown in this report card are simulated and for demonstration only**. 


```{r setup, echo = F, include = F}
library(ggplot2)
library(plotly)
library(readxl)
library(dplyr)
library(sf)
library(colorspace)
library(ggsflabel)

set.seed(07032025)

# Set your own working directory
kelp_data_folder <- file.path("~", "..", "ESSA Dropbox","d_global","Projects","EN2900to2999",
  "EN2955 - MaPP Kelp State of Knowledge","Data",
  "MaPP Site Metadata")
dir(kelp_data_folder)
```

:::: {style='display: flex; justify-content: flex-start; column-gap: 10px;'}

::: {}

```{r map it up, echo = F, fig.out = c(8, 8), message = F, warning = F, dpi = 150}
t2_sites <- read.csv(file.path(kelp_data_folder, 
  "Subregional_T2_DataSummary.csv"))
t2_sites_sf <- st_transform(st_as_sf(t2_sites, coords = c("Long", "Lat"), 
      crs = 4326),
    crs = 3005)

t2_sites_sf <- t2_sites_sf |>
  mutate(Subregion2 = case_when(
    Subregion %in% c("KXX", "Heiltsuk") ~ "Central Coast",
    Subregion == 'NVI' ~ "North Vancouver Isl.",
    TRUE ~ Subregion
  ))

# Create a bounding box by Subregion2
st_bbox_by_feature = function(x, subregionID) {
  y = st_geometry(x |> filter(Subregion2 == subregionID))
  st_as_sfc(st_bbox(y))
}

# Generate fake data; what is the status of each bed?
status_vec <- c("Increasing (+20% or more)", "No trend", "Decreasing (-20% or more)", # "Total bed loss", 
"Unknown")

t2_sites_sf <- t2_sites_sf |>
  group_by(Subregion2) |>
  mutate(n = n()) |>
  ungroup() |>
  mutate(example_status = sample(status_vec, n(), replace = TRUE))
t2_sites_sf$example_status[which(t2_sites_sf$n < 2)] <- NA

# Function to get the "mode" of a character vector
mode_char <- function(x) unique(x)[which.max(tabulate(match(x, unique(x))))]

bboxes <- list()
t2_sites_sf$regionstatus <- NA
for(i in unique(t2_sites_sf$Subregion2)) {
  ex_regionstatus <- mode_char(t2_sites_sf |> filter(Subregion2 == i) |> pull(example_status))
  bboxes[[i]] <- st_as_sf(st_bbox_by_feature(x = t2_sites_sf, subregionID = i)) |>
    mutate(regionstatus = ex_regionstatus)
  t2_sites_sf$regionstatus[which(t2_sites_sf$Subregion2 == i)] <- ex_regionstatus
        # st_transform(crs = 3005) |>
        # st_difference(bc_highres)
}

globalbbox <- st_bbox(t2_sites_sf)

kelppalette <- colorspace::diverge_hcl(5, palette = "Red-Green")

(pl <- ggplot(data = bboxes[[1]]) + 
  geom_sf(data = bboxes[[1]], aes(fill = regionstatus), alpha = 0.8, color = NA) + 
  geom_sf(data = bboxes[[2]], aes(fill = regionstatus), alpha = 0.8, color = NA) +
  geom_sf(data = bboxes[[3]], aes(fill = regionstatus), alpha = 0.8, color = NA) + 
  geom_sf(data = bboxes[[4]], aes(fill = regionstatus), alpha = 0.8, color = NA) + 
  # geom_sf(data = bboxes[[5]], fill = mappalette[5], alpha = 0.8, color = NA) + 
  geom_sf(data = bcmaps::bc_bound_hres() |> st_transform(crs = 3005),
    fill = "grey80", color = "grey80") + 
  geom_sf_label_repel(data = t2_sites_sf |> group_by(Subregion2) |> slice(1), 
    aes(label = Subregion2), fill = alpha("white", 0.5)) + 
  scale_fill_manual(name = "Sub-regional status", 
    values = c(kelppalette[5], kelppalette[4], kelppalette[2], "slategrey"),
    drop = FALSE,
    breaks = status_vec,
    labels = status_vec) +
  theme_minimal() + 
  # theme(
  #   legend.position = "bottom",
  #   legend.box="vertical",
  #   legend.direction = "vertical"
  # ) + 
  coord_sf(
    xlim = c(globalbbox$xmin - 200, globalbbox$xmax + 200), 
    ylim = c(globalbbox$ymin - 200, globalbbox$ymax + 200))
  ) + 
  labs(y = "Longitude", x = "Latitude") +
  theme(legend.position = "none")
```

:::

::::

:::: {style='display: flex; justify-content: flex-start; column-gap: 10px;'}

::: {}


```{r create_map, echo = F, warning = F, fig.out = c(8, 8), dpi = 150}

ggplot(t2_sites_sf, aes(x = factor(Subregion2, levels = c("North Coast", "Haida Gwaii", "Central Coast", "North Vancouver Isl.")), fill = example_status)) + 
  geom_bar(position = "fill") +
  scale_fill_manual(name = "Sub-regional status", 
    values = c(kelppalette[5], kelppalette[4], kelppalette[2], "slategrey"),
    drop = FALSE,
    breaks = status_vec,
    labels = status_vec) +
  theme_minimal() +
  labs(y = "Status", x = "Sub-region") +
  geom_text(data = t2_sites_sf |> group_by(Subregion2) |> slice(1), 
    aes(label = paste0(n, " sites")), y = 1.05) + 
  expand_limits(y = 1.1) + 
  # theme(
  #   legend.position = "bottom",
  #   legend.box="vertical",
  #   legend.direction = "vertical"
  # ) + 
  scale_y_continuous(breaks = c(0.25, 0.5, 0.75, 1))
  # facet_wrap(~Subregion2, scales = "free_x")
```

::: 


::::

**Summary of kelp bed area status**. Each site within a subregion has its own status; the sub-region status is the mode of the individual site statuses. Above each bar in the plot above is the number of sites in that sub-region.
